Concurrent Programming with CXL

### 2020-06-19 book/paper.tex
Latest

\begin{figure}
\begin{code}
\begin{verbatim}
   ctx = __init__/() pc =   0 #states = 1 diameter = 0 queue = 0
CXL Assertion failed ('assert', 'Up.cxl', 9, 5) 1
#states = 79
==== Safety violation ====
__init__/() [0,32-47]       48 dict{ .count:0, .done:dict{ 0:False, 1:False } }
incrementer/0 [1-4]          5 dict{ .count:0, .done:dict{ 0:False, 1:False } }
incrementer/1 [1-8]          9 dict{ .count:1, .done:dict{ 0:False, 1:True  } }
incrementer/0 [5-8]          9 dict{ .count:1, .done:dict{ 0:True,  1:True  } }
main/() [11-15,18-22,24-29] 29 dict{ .count:1, .done:dict{ 0:True,  1:True  } }
\end{verbatim}
\end{code}
\caption{The output of running Figure~\ref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

@@

The state of a process is called a \emph{context} (aka continuation):
\index{context}
\index{continuation}
\glsadd{context}
it contains the value of
its name tag, program counter, stack, and register.
The state of a CVM
consists of the value of its memory and the multiset (or \emph{bag})
\index{bag}
\index{multiset}
of contexts.  It is a multiset of contexts because two different processes can
be in the same state.
The initial state of the CXL memory is the empty dictionary, \texttt{dict\{\}}.
The context bag has an initial context in it with name tag
\texttt{\_\_main\_\_/()}, pc 0, register \texttt{()}, and an empty stack.
Each machine instruction updates the state in some way.
Figure~\ref{fig:incstate} shows an example of a reachable state for 
the program in Figure~\ref{fig:inc}.

@@

Compiling the code in Figure~\ref{fig:inc} results in the CVM bytecode
listed in Figure~\ref{fig:inccode}.
You can obtain this code by invoking \texttt{cxl} with the \texttt{-a} flag
like so:
\begin{code}
\begin{verbatim}
cxl -a Up.cxl
\end{verbatim}
\end{code}
Each process in the CVM is predominantly a \emph{stack machine},
\index{stack machine}
\glsadd{stack machine}
but it also has a register.
All instructions are atomically executed.
Most instructions pop values from the stack or push values onto the stack.
At first there is one process (with name tag \texttt{\_\_init\_\_/()}
that initializes the state.
It starts executing at instruction 0 and keeps executing until the last
execution in the program.
In this case, the first instruction is a \texttt{JUMP} instruction that sets the
program counter to 39.
At program counter 1 is the code for the \texttt{incrementer} method.
All methods start with a \texttt{Frame} instruction and end with a \texttt{Return}
instruction.

Appendix~\ref{ap:cxlbytecode} provides a list of all CVM machine instructions,
in case you want to read about the details.
The \texttt{Frame} instruction lists the name of the method,
names of its arguments, and
the program counter of the method's \texttt{Return} instruction.
The code generated from \texttt{count := count + 1} in line 2 of
\texttt{Up.cxl} is as follows:

>>>>

\begin{figure}
\begin{code}
\begin{verbatim}
#states = 79
==== Safety violation ====
CXL Assertion failed ('assert', 'Up.cxl', 9, 5) 1
__init__/() [0,32-47]       48 dict{ .count:0, .done:dict{ 0:False, 1:False } }
incrementer/0 [1-4]          5 dict{ .count:0, .done:dict{ 0:False, 1:False } }
incrementer/1 [1-8]          9 dict{ .count:1, .done:dict{ 0:False, 1:True  } }
incrementer/0 [5-8]          9 dict{ .count:1, .done:dict{ 0:True,  1:True  } }
main/() [11-15,18-22,24-29] 29 dict{ .count:1, .done:dict{ 0:True,  1:True  } }
\end{verbatim}
\end{code}
\caption{The output of running Figure~\ref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

@@

The state of a process is called a \emph{context} (aka \emph{continuation}):
\index{context}
\index{continuation}
\glsadd{context}
it contains the value of
its name tag, program counter, stack, and register.
The state of a CVM
consists of the value of its memory and the multiset (or \emph{bag})
\index{bag}
\index{multiset}
of contexts.  It is a multiset of contexts because two different processes can
be in the same state.
The initial state of the CXL memory is the empty dictionary, \texttt{dict\{\}}.
The context bag has an initial context in it with name tag
\texttt{\_\_main\_\_/()}, pc 0, register \texttt{()}, and an empty stack.
Each machine instruction updates the state in some way.
% Figure~\ref{fig:incstate} shows an example of a reachable state for the program in Figure~\ref{fig:inc}.

@@

Compiling the code in Figure~\ref{fig:inc} results in the CVM bytecode
listed in Figure~\ref{fig:inccode}.
You can obtain this code by invoking \texttt{cxl} with the \texttt{-a} flag
like so:
\begin{code}
\begin{verbatim}
cxl -a Up.cxl
\end{verbatim}
\end{code}
Each process in the CVM is predominantly a \emph{stack machine},
\index{stack machine}
\glsadd{stack machine}
but it also has a register.
All instructions are atomically executed.
Most instructions pop values from the stack or push values onto the stack.
At first there is one process, with name tag \texttt{\_\_init\_\_/()},
that initializes the state.
It starts executing at instruction 0 and keeps executing until the last
execution in the program.
In this case, the first instruction is a \texttt{JUMP} instruction that sets the
program counter to 39.
At program counter 1 is the code for the \texttt{incrementer} method.
All methods start with a \texttt{Frame} instruction and end with a \texttt{Return}
instruction.

Appendix~\ref{ap:cxlbytecode} provides a list of all CVM machine instructions,
in case you want to read about the details.
The \texttt{Frame} instruction lists the name of the method,
names of its arguments, and
the program counter of the method's \texttt{Return} instruction.
The code generated from \texttt{count := count + 1} in line 2 of
\texttt{Up.cxl} is as follows (see Figure~\ref{fig:inccode}):
